---
title: "Example_MTS_VAR1_MBSG"
output: html_document
---

February 1, 2022

This is an example notebook for generating MBSG influence and vulnerability scores to identify source anddd sink nodes with simulated VAR(1) multivariate time series under DCG network structure and independent Gaussian errors. Comparisons with Standard VAR is also included.

##Load Packages
```{r, warning=FALSE, message=FALSE}
library(MTS)
library(mvtnorm)
library(knitr)
library(data.table)
library(cIRT)
library(fBasics)
library(MASS)
library(igraph)
```

##Load MBSG Functions
```{r}
#Functions for Posterior Generation: works for any VAR(p)
#Generate M posterior samples for 1 realization of time series
design.matrix.from.ts <- function(zt, p) {
nT <- nrow(zt)
x = rep(1, nT-p)
for (i in 1:p){
  x = cbind(x, zt[(p+1-i):(nT-i), ])
}
return(as.matrix(x))
}
gen_posterior <- function(zt, p = 1, lambda=10, V0=NULL, n0=5, nsamples=1000, intercept=FALSE){
  X = design.matrix.from.ts(zt, p)[, -1] #extracting design matrix from ts
  k = ncol(zt) #dimension of ts
  nT = nrow(zt) #total T obs
  
  #Setting up initial matrices
  C = solve(lambda*diag(k*p)) #precision matrix of coefficient matrix
  B0 = matrix(0, k*p, k)
  if (intercept){
    C = solve(lambda*diag(k*p+1)) #precision matrix of coefficient matrix
    B0 = matrix(0, k*p+1, k) #hyperparameter, prior mean for parameters
    X = design.matrix.from.ts(zt, p) #extracting design matrix from ts
  }
  if (is.null(V0)){
    V0 = diag(k) # k-by-k covariance matrix for prior of Sigma_a
  }
  
  ne = nT-p
  Z = zt[(1+p):nT, ] #(T-p)-by-k matrix
  Bhat = solve(t(X)%*%X) %*% t(X) %*%Z #LS estimate
  Btilde = solve(t(X)%*%X + C) %*% (t(X) %*% X %*% Bhat + C %*% B0) #posterior mean 
  Stilde = t(Z - X%*%Btilde) %*% (Z-X%*%Btilde) + t(Btilde- B0)%*% C %*% (Btilde - B0)
  XXC = solve(t(X)%*%X + C)
  V0S = V0 + Stilde
  
  pos_Beta = as.list(1:nsamples)
  pos_BetaNoIntercept = as.list(1:nsamples)
  pos_Sigma <- as.list(1:nsamples)
  
  num_columns = k*p
  if (intercept){
    num_columns = k*p+1
  }
  
  set.seed(2020)
  for (i in 1:nsamples){
  Sigma_a <- riwishart(n0+ne, V0S) #This gives an kxk matrix
  pos_Sigma[[i]] <- Sigma_a
  beta_sample <- mvrnorm(1, vec(Btilde), Sigma_a %x% XXC) #row stack of kx(kp+1) matrix if intercept is TRUE
  beta_matrix <- matrix(beta_sample, nrow=k, ncol=num_columns, byrow = TRUE) #convert back to matrix
  
  pos_Beta[[i]] <- beta_matrix
  pos_BetaNoIntercept[[i]] <- beta_matrix
  if (intercept){
    pos_BetaNoIntercept[[i]] <- beta_matrix[, -1] #Remove the first column of intercepts
  }
  }
  
  #Getting point estimates from built-in BVAR
  phi0=matrix(0, nrow=num_columns, ncol = num_columns) #BVAR does not run when intercept = FALSE and Phi0 is set to default NULL. Must add phi0 manually. Phi0 has an issue when include.mean=FALSE.
  point_est = NA
  if (intercept){
    point_est =  BVAR(zt, p=p, C=lambda*diag(num_columns),V0=diag(k), 
                   include.mean = intercept, Phi0 = NULL)
  }else{
    point_est = BVAR(zt, p=p, C=lambda*diag(num_columns),V0=diag(k), 
                   include.mean = intercept, Phi0 = phi0)
  }
  point_estBeta = point_est$Phi
  if(intercept){
     point_estBeta = cbind(point_est$phi0, point_est$Phi)
  }
  point_estSigma = point_est$Sigma
  return(list(pos_Beta = pos_Beta, pos_Sigma = pos_Sigma, 
              pos_BetaNoIntercept = pos_BetaNoIntercept,
              point_estBeta = point_estBeta, point_estSigma=point_estSigma, 
              point_est = point_est))
}
gen_spillovers <- function(pos_Beta, pos_Sigma, lag, intercept=FALSE){
  nsamples = length(pos_Beta)
  k = nrow(pos_Sigma[[1]])
  pos_spillover = as.list(1:nsamples)
  pos_nspillover = as.list(1:nsamples)
  pos_spilloverindex = rep(NA, times=nsamples)
  
  set.seed(2020)
  for (i in 1:nsamples){
    pos_Beta_sample = pos_Beta[[i]]
    if (intercept){
      pos_Beta_sample = pos_Beta_sample[, -1]
    }
    fevd_obj = FEVdec(pos_Beta_sample, Theta = NULL, pos_Sigma[[i]], lag=lag) 
    start = lag*k + 1
    end = (lag+1) * k
    w =  fevd_obj$Omega[, start:end]
    w_norm = fevd_obj$OmegaR[, start:end] * 100
    pos_spillover[[i]] <- w
    pos_nspillover[[i]] <- w_norm
    pos_spilloverindex[i] <-  sum(w_norm) - tr(w_norm)
    
  }
  return(list(pos_spillover = pos_spillover, pos_nspillover = pos_nspillover, pos_spilloverindex = pos_spilloverindex))
}

extract_pd <- function(input, i, j){
  #input should be list of posterior sample matrices
  pos_Beta <- input
  posvalues <- rep(NA, length(pos_Beta))
  for (d in 1:length(pos_Beta)){
    posvalues[d] <- pos_Beta[[d]][i, j]
  }
  return(posvalues)
}
extract_scores <- function(nspillover){
  nsamples = length(nspillover)
  ncomponents = ncol(nspillover[[1]])
  importance_score = as.list(1:nsamples)
  vulnerability_score = as.list(1:nsamples)
  for (i in 1:nsamples){
    w = nspillover[[i]]
    toSubtract = diag(nrow(w))
    diag(toSubtract) <- diag(w)
    s = w-toSubtract
    importance = apply(s, 2, sum) #Column sums of all impact caused by node k
    importance_score[[i]] = (importance/sum(s)) * 100
    vulnerability_score[[i]] = apply(s, 1, sum) #Row sums of all impact going into node j
  }
  importance_score = do.call(rbind, importance_score)
  vulnerability_score = do.call(rbind, vulnerability_score)
  return(list(importance_score = importance_score, vulnerability_score = vulnerability_score))
}

#Load NDCG computations
NDCG_calc = function(pred_ranking, ground_truth_order, ground_truth_rel, large2small = TRUE){
  print(pred_ranking)
  pred_order = order(pred_ranking, decreasing=large2small)
  print(pred_order)
  DCG = sum(ground_truth_rel/log2(pred_order + 1))
  print(DCG)
  IDCG = sum(ground_truth_rel/log2(ground_truth_order + 1))
  print(IDCG)
  NDCG = DCG/IDCG
  return(NDCG)
}
```


##Load Data
```{r}
MTSreps = list.files(".", pattern = "*MTSrep")

MTS = read.csv(MTSreps[1])
plot.ts(MTS[, 1:10])
```

## Run MBSG for 1:h
```{r}
#Specify ground truth
d=20; nsink=5; nsource=5; nmedium=d-nsink-nsource; h = 10

#Report accuracy (NDCG) for source/sink nodes based on MBSG influence and vulnerability. 
imp_ndcg = as.list(1:5)
vul_ndcg = as.list(1:5)

for (j in 1:5){
  zt = read.csv(MTSreps[j])
  zt = as.matrix(zt)
  bfitted = gen_posterior(zt, p=1, lambda=10, V0=cov(zt), n0=5, nsamples=100, intercept = FALSE)
  
  #importance scores
  ground_truth_order = 1:dd
  truth_rel_imp = c(rep(1, nsource), rep(0.5, nmedium), rep(0, nsink))

  #vulnerability scores
  ground_truth_order_vul = d:1
  truth_rel_vul = c(rep(0, nsource), rep(0.5, nmedium), rep(1, nsink))
  
  ###Record bsg estimates for each h
  imp_ndcg_h = rep(NA, times=h)
  vul_ndcg_h = rep(NA, times=h)

  ###Generate h forecasts
  for (i in 1:h){
    output_spillover <- gen_spillovers(bfitted$pos_Beta, bfitted$pos_Sigma, lag=i, intercept = FALSE)
    output_scores <-  extract_scores(output_spillover$pos_nspillover)
    importanceDF = data.frame(output_scores[[1]])
    vulnerabilityDF = data.frame(output_scores[[2]])
    bsg_imp_ranking = colMeans(importanceDF)
    bsg_imp_ndcg = NDCG_calc(bsg_imp_ranking, ground_truth_order, truth_rel_imp)
    
    bsg_vul_ranking = colMeans(vulnerabilityDF) 
    bsg_vul_ndcg = NDCG_calc(bsg_vul_ranking, ground_truth_order_vul, truth_rel_vul, large2small = TRUE)
    
    imp_ndcg_h[i] <- bsg_imp_ndcg
    vul_ndcg_h[i] <- bsg_vul_ndcg
    print(paste("Done with BSG iter:", h))
  }
  imp_ndcg[[j]] <- imp_ndcg_h
  vul_ndcg[[j]] <- vul_ndcg_h
}


impDF = do.call(rbind, imp_ndcg)
vulDF = do.call(rbind, vul_ndcg)
DF = data.frame(rbind(impDF, vulDF))
colnames(DF) = paste("MBSG:h=", 1:h, sep="")
DF$Score = rep(c("Importance", "Vulnerability"), each=5)

DF #NDCG for each of 5 replicates
```

## Run Standard VAR with FDR control
```{r}
#Set ground truth
d=20; nsource=5; nsink=5; nmedium=d-nsource-nsink
imp_ndcg = as.list(1:5)
vul_ndcg = as.list(1:5)

for (i in 1:5){
  zt = read.csv(MTSreps[i])
  #Baseline VAR Network Granger Causality w/ shrinkage from FDR
  fitted_VAR = VAR(zt, p=1, include.mean = FALSE)

  networkgc = fitted_VAR$Phi
  gc_pvals = 1- pnorm(abs(fitted_VAR$Phi/fitted_VAR$Sigma)) #Two sided test
  gc_pvals2 = c(gc_pvals) #This unlists by columns, e.g. c1, c2, c3
  gc_pvals_adj = matrix(p.adjust(c(gc_pvals), method="fdr"), nrow=20, ncol=20, byrow = FALSE)
  networkgc_adj = networkgc
  networkgc_adj[gc_pvals_adj > 0.05] <- 0

  est_graph <-  graph_from_adjacency_matrix(t(networkgc_adj), mode = "directed", weighted = TRUE, diag=FALSE)
  plot(est_graph, main="Standard VAR")

  #importance scores
  ground_truth_order = 1:d
  truth_rel_imp = c(rep(1, nsource), rep(0.5, nmedium), rep(0, nsink))
  
  outdeg_ranking = degree(est_graph, normalized=TRUE, mode = "out")
  outdeg_imp_ndcg = NDCG_calc(outdeg_ranking, ground_truth_order, truth_rel_imp, large2small=TRUE)
  
  eigen_ranking = eigen_centrality(est_graph, scale=TRUE)$vector
  eigen_imp_ndcg = NDCG_calc(eigen_ranking, ground_truth_order, truth_rel_imp, large2small=TRUE)
  
  between_ranking=betweenness(est_graph, weights=abs(E(est_graph)$weight))
  between_imp_ndcg = NDCG_calc(between_ranking, ground_truth_order, truth_rel_imp, large2small=TRUE)
  
  closeness_ranking=closeness(est_graph, weights=abs(E(est_graph)$weight))
  closeness_imp_ndcg = NDCG_calc(closeness_ranking, ground_truth_order, truth_rel_imp, large2small=TRUE)
  
  #vulnerability scores
  ground_truth_order_vul = d:1
  truth_rel_vul = c(rep(0, nsource), rep(0.5, nmedium), rep(1, nsink))
  
  eigen_ranking = eigen_centrality(est_graph, scale=TRUE)$vector
  eigen_vul_ndcg = NDCG_calc(eigen_ranking, ground_truth_order_vul, truth_rel_vul, large2small = TRUE)
  
  between_ranking=betweenness(est_graph, weights=abs(E(est_graph)$weight))
  between_vul_ndcg = NDCG_calc(between_ranking,  ground_truth_order_vul, truth_rel_vul, large2small = TRUE)
  
  closeness_ranking=closeness(est_graph, weights=abs(E(est_graph)$weight))
  closeness_vul_ndcg = NDCG_calc(closeness_ranking,  ground_truth_order_vul, truth_rel_vul, large2small = TRUE)
  
  indeg_ranking = degree(est_graph, normalized=TRUE, mode="in") #high in degree is more vulnerable
  indeg_vul_ndcg = NDCG_calc(indeg_ranking, ground_truth_order_vul, truth_rel_vul, large2small = TRUE)
  
  imp_ndcg[[i]] <- data.frame(Degree = outdeg_imp_ndcg, Eigen = eigen_imp_ndcg, 
                        Between = between_imp_ndcg, Closeness = closeness_imp_ndcg)
  
  vul_ndcg[[i]] <- data.frame(Degree = indeg_vul_ndcg, Eigen = eigen_vul_ndcg, 
                        Between = between_vul_ndcg, Closeness = closeness_vul_ndcg)

  print(paste("Done with iteration", i))
}

impDF = do.call(rbind, imp_ndcg)
vulDF = do.call(rbind, vul_ndcg)
DF = data.frame(rbind(impDF, vulDF))
DF$Score = rep(c("Importance", "Vulnerability"), each=5)

DF #NDCG


```

